---
title: "Retrieve Data using an API"
author: "Felicia Cruz"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(metajam)
```
We are using `metajam` to download HAB data for the CA coast from DataOne. 

"California Harmful Algal bloom Monitoring and ALert Program Data"

https://search.dataone.org/view/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fmetadata%2Feml%2Fedi%2F988%2F4#https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F4119639092e62c55ea8be348e4d9260d

## Event data download
- By looking at the metadata, this includes location and geometric info 
```{r, error=TRUE}
event_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F4119639092e62c55ea8be348e4d9260d"

path <- here()

event_data <- download_d1_data(event_data_url, path)

# returns a list of all of the data in the folder 
all_event_data <- metajam::read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__event__csv"))

View(all_event_data$data)
```

## Occurrence Data
- From looking at the metadata, this includes organism information 
```{r, error = TRUE}
occ_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F7ad56af78ac5f73d81f5e80293f1d6c9"

occ_data <- download_d1_data(occ_data_url, path)

all_occ_data <- read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__occurrence__csv"))

View(all_occ_data$attribute_metadata)
```

## Extended Measurement Data 
- From the metadata, this contains information about measurement type, value, etc.  
```{r, error = TRUE}
ext_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2Ff616903aa9a2592cb3082cb3b9525270"

ext_data <- download_d1_data(ext_data_url, path)

all_ext_data <- read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__extendedmeasurementorfact__csv"))

View(all_ext_data$attribute_metadata)
View(all_ext_data$data)
```

