---
title: "Retrieve Data using an API"
author: "Felicia Cruz"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(metajam)
library(lubridate)

```
We are using `metajam` to download HAB data for the CA coast from DataOne. We wil specifically filter for locations at the Scripps Pier in La Jolla.

"California Harmful Algal bloom Monitoring and ALert Program Data"

https://search.dataone.org/view/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fmetadata%2Feml%2Fedi%2F988%2F4#https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F4119639092e62c55ea8be348e4d9260d

https://cran.r-project.org/web/packages/rnoaa/rnoaa.pdf

```{r}
# set file path
path <- here()

```


## Data download: Event, Occurrence, and Extended Measurement Data.
 - Only run if data not already downloaded locally
```{r, eval=FALSE}

# Event Data Download
event_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F4119639092e62c55ea8be348e4d9260d"
event_data <- download_d1_data(event_data_url, path)


# Occurrence Data Download
occ_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F7ad56af78ac5f73d81f5e80293f1d6c9"
occ_data <- download_d1_data(occ_data_url, path)


# Extended Measurement Download
ext_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2Ff616903aa9a2592cb3082cb3b9525270"
ext_data <- download_d1_data(ext_data_url, path)
```

## 
```{r}

# returns a list of all of the data in the folder 
all_event_data <- metajam::read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__event__csv"))

all_occ_data <- read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__occurrence__csv"))

all_ext_data <- read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__extendedmeasurementorfact__csv"))


```


## Event data filtering
- By looking at the metadata, this includes location and geometric info 
```{r, error=TRUE}

# view all event data
View(all_event_data$data)

# filter for ScrippsPier
scripps <- all_event_data$data %>% 
  filter(locationID == "HABs-ScrippsPier") %>% 
  data.frame()

view(scripps)

occ <- all_occ_data$data %>% data.frame()
ext <- all_ext_data$data %>% data.frame()

```

# join different data sets together.  
```{r}
# joins Scripps event data and occurrence data
scripps_1 <- left_join(scripps, occ, by = "id")

# joins Scripps event data/occurrence data with extended measurement data
scripps_combined <- left_join(scripps_1, ext, by = "id")
```


```{r}

unique(scripps_combined$measurementType)


```



## Occurrence Data
- From looking at the data and metadata, this includes organism information 
```{r, error = TRUE}

View(all_occ_data$attribute_metadata)
view(all_occ_data$data)
```

## Extended Measurement Data 
- From the data and metadata, this contains information about measurement type, value, etc.  
```{r, error = TRUE}

View(all_ext_data$attribute_metadata)
View(all_ext_data$data)
```

