---
title: "Retrieve Data using an API"
author: "Felicia Cruz"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(metajam)
library(sf)
```
We are using `metajam` to download HAB data for the CA coast from DataOne. We wil specifically filter for locations at the Scripps Pier in La Jolla.

"California Harmful Algal bloom Monitoring and ALert Program Data"

https://search.dataone.org/view/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fmetadata%2Feml%2Fedi%2F988%2F4#https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F4119639092e62c55ea8be348e4d9260d



## All data download: Event, Occurrence, and Extended Measurement Data
```{r, eval=FALSE}
path <- here()

# Event Data Download
event_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F4119639092e62c55ea8be348e4d9260d"

event_data <- download_d1_data(event_data_url, path)

# returns a list of all of the data in the folder 
all_event_data <- metajam::read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__event__csv"))


# Occurrence Data Download
occ_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2F7ad56af78ac5f73d81f5e80293f1d6c9"
occ_data <- download_d1_data(occ_data_url, path)

all_occ_data <- read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__occurrence__csv"))


ext_data_url <- "https://cn.dataone.org/cn/v2/resolve/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fdata%2Feml%2Fedi%2F988%2F4%2Ff616903aa9a2592cb3082cb3b9525270"

ext_data <- download_d1_data(ext_data_url, path)

all_ext_data <- read_d1_files(here("https_pasta.lternet.edu_package_metadata_eml_edi_988_4__extendedmeasurementorfact__csv"))

```


## Event data filtering
- By looking at the metadata, this includes location and geometric info 
```{r, error=TRUE}

# view all event data
View(all_event_data$data)

# filter for ScrippsPier
scripps_events <- all_event_data$data %>% 
  filter(locationID == "HABs-ScrippsPier")

view(scripps_events)

```

```{r}
ggplot(data = scripps_events, aes(x = eventDate)) +
  geom_hist()

```



## Occurrence Data
- From looking at the data and metadata, this includes organism information 
```{r, error = TRUE}

View(all_occ_data$attribute_metadata)
view(all_occ_data$data)
```

## Extended Measurement Data 
- From the data and metadata, this contains information about measurement type, value, etc.  
```{r, error = TRUE}

View(all_ext_data$attribute_metadata)
View(all_ext_data$data)
```

